---
title: "Code for 'Pandemic Pivots for Older Volunteers: Online Tutoring and Pen Pal Programs'"
author: "Peter Sun"
date: "September, 2022"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_depth: 3
---

# Import Packages

```{r warning=F, error=F, message=F}
library(tidyverse)
library(knitr)
library(kableExtra)
library(readxl)
library(glue)
library(viridis)
library(lubridate)
library(ggsci)
library(scales)
library(ggthemes)
library(kableExtra)
library(lsr)
library(chisq.posthoc.test)
library(tableone)
library(mice)
```

# Import Data and Labels

```{r warning=F, error=F, message=F}
# Import the following data:
# (1) df =  the study sample (respondents who completed both pre and post)
# (2) df.pre.only = respondents who only completed pre
# (3) df.codebook = labels for analytical variables
# (4) df.imputed = 55 imputations using mice
load("private/data.Rdata")
```

# Inspect Variables

This dataset contains 61 rows (the study sample size) and 68 columns/variables.

```{r warning=F, error=F, message=F}
dim(df)
```

The variables are categorized as follows:

- 3 administrative variables (prefixed *admin_*)
- 20 recoded variables for bivariate analyses (prefixed *analysis_*)
- 4 variables corresponding to open-ended questions (prefixed *openended_*)
- 14 original pre-test and post-test variables prior to recoding (prefixed *original_*)
- 9 pre-test survey demographic variables for Table 1 (prefixed *tab1_*)
- 11 post-test survey outcome variables for Table 2 (prefixed *tab2_*)
- 7 variables from the parent study as multiple imputation predictors (prefixed *imp_*)

```{r warning=F, error=F, message=F}
df %>%
  names() %>%
  str_extract("[:alnum:]+") %>%
  enframe(value = "variable_type") %>%
  count(variable_type) %>%
  janitor::adorn_totals()
```

The 20 variables used for the quantitative bivariate analyses are as follows:

```{r}
df %>%
  select(starts_with("analysis_")) %>%
  names()
```

# Missing Data

Missing data for the 20 variables (for bivariate analyses) in this study range from 1.6% (1 out of 61 responses missing) to 37.7% (23 out of 61 responses missing).

```{r warning=F, error=F, message=F}
df %>%
  select(admin_id, starts_with("analysis")) %>%
  mutate(across(-admin_id, ~is.na(.))) %>%
  pivot_longer(-admin_id) %>%
  count(name, value) %>%
  pivot_wider(names_from = "value", values_from = "n") %>%
  setNames(c("variable", "nonmissing", "missing")) %>%
  mutate(pct_missing = missing/(nonmissing+missing)) %>%
  arrange(desc(pct_missing)) %>%
  mutate(n = row_number()) %>%
  relocate(n) %>%
  mutate(pct_missing = percent(pct_missing)) %>%
  kbl() %>%
  kable_paper()
```

For the bivariate models, missing data ranges from 6.6% to 54.1%:

```{r warning=F, error=F, message=F}
df.tmp <- tibble(IV = character(0), DV = character(0), pct_missing = numeric(0))
iv <- df %>% select(starts_with("analysis_pre_")) %>% names()
dv <- df %>% select(starts_with("analysis_post_")) %>% names()
for(d in dv) {
  for(i in iv) {
    n <- df %>%
      select(i, d) %>%
      drop_na() %>%
      nrow()
    df.tmp <- df.tmp%>%
      add_row(IV = i, DV = d, pct_missing = (61 - n) / 61)
  }
}
mean(df.tmp$pct_missing) # Mean % Missing
df.tmp %>%
  arrange(desc(pct_missing)) %>%
  mutate(pct_missing = percent(pct_missing, accuracy = 0.1)) %>%
  kbl() %>%
  kable_paper()
```

# Multiple Imputation

Therefore, 55 imputations is appropriate for this study.

```{r}
# # Select study variables for imputation
# # And drop empty factor levels for mice to work
# df.for.imputation <- df %>%
#   select(starts_with("analysis"), starts_with("imp")) %>%
#   mutate(across(where(is.factor), ~fct_drop(.)))
# names(df.for.imputation)
# length(df.for.imputation)
# 
# # Multiple Imputation of Study Variables
# df.imputed = mice(df.for.imputation, m = 55, maxit = 25, print = F)
# 
# # Save Imputed Data
# saveRDS(df.imputed, "data_imputed.Rds")
```

# Table 1

```{r warning=F, error=F, message=F}
# Select Table 1 Variables and Labels
df.table1 <- df %>% 
  select(admin_program, starts_with("tab1_")) %>%
  setNames(c("Program", df.codebook %>% 
                          filter(str_detect(variable, "tab1_")) %>% pull(label)))

# Demographics by Program
table1 <- CreateTableOne(data = df.table1, strata = "Program", test = T)
table1.print <- print(table1, quote = F, noSpaces = T, printToggle = F, smd = F, pDigits = 2)
table1.print %>% kbl() %>% kable_paper()
# write.csv(table1.print, "table1.csv")

# Totals
table1.totals <- CreateTableOne(data = df.table1, test = F)
table1.totals.print <- print(table1.totals, quote = F, noSpaces = T, printToggle = F, smd = F, pDigits = 2)
table1.totals.print %>% kbl() %>% kable_paper()
# write.csv(table1.totals.print, "table1.totals.csv")
```

# Table 2

```{r warning=F, error=F, message=F}
# Select Table 2 Variables and Labels
df.table2 <- df %>% 
  select(admin_program, starts_with("tab2_")) %>%
  setNames(c("Program", df.codebook %>% 
                          filter(str_detect(variable, "tab2_")) %>% pull(label)))
  
# Table 2 By Program
table2 <- CreateTableOne(data = df.table2, strata = "Program", test = T)
table2.print <- print(table2, quote = F, noSpaces = T, printToggle = F, smd = F, pDigits = 2)
table2.print %>% kbl() %>% kable_paper()
# write.csv(table2.print, "table2.csv")

# Table 2 Totals
table2.totals <- CreateTableOne(data = df.table2, test = F)
table2.totals.print <- print(table2.totals, quote = F, noSpaces = T, printToggle = F, smd = F, pDigits = 2)
table2.totals.print %>% kbl() %>% kable_paper()
# write.csv(table2.totals.print, "table2.totals.csv")
```

# Functions for Bivariate Tests

```{r warning=F, error=F, message=F}
# Function for chi-square test, Cramer's V, Bonferroni post-hoc test, and bar plot
easy_chi <- function(x_var, y_var, data) {

  # Convert strings into symbols
  x_var <- sym(x_var)
  y_var <- sym(y_var)
  
  # Chi-square test and Cramer's V
  cs <- chisq.test(pull(data, x_var), pull(data, y_var)) %>%
    tidy() %>%
    mutate(CramersV = lsr::cramersV(
      pull(data, x_var), pull(data, y_var)))
  
  # Build caption
  caption <- paste0("chi = ", round(cs$statistic, 2), "; p = ", 
                    round(cs$p.value, 2), "; Cramer's V = ", 
                    round(cs$CramersV, 2))
  
  # Calculate frequencies
  freqs <- data %>%
    select({{ x_var }}, {{ y_var }}) %>%
    drop_na() %>%
    count({{ x_var }}, {{ y_var }}) %>%
    group_by({{ x_var }}) %>%
    mutate(pct = n/sum(n)) %>%
    ungroup() %>%
    mutate(text = paste0(n, " (", percent(pct, accuracy = 0.1), ")"))
  
  # Plot
  plot <- ggplot(freqs, aes(x = {{ x_var }}, y = pct, fill = {{ y_var }})) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(aes(label = text, group = {{ y_var }}), 
      position = position_stack(vjust = 0.5),
      size = 2.5,
      color = "gray30") +
    theme_classic() +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_npg() +
    labs(caption = caption) +
    theme(legend.position = "top",
    plot.title = element_text(size = 18),
    plot.caption = element_text(size = 10),
    text = element_text(size = 10))
  
  # Bonferroni post-hoc test
  ph <- chisq.posthoc.test::chisq.posthoc.test(
    table(select(data, {{ x_var }}, {{ y_var }})))
  
  # Return objects
  return(list(plot = plot, freqs = freqs, ph = ph, 
              chi = cs$statistic, chi.p = cs$p.value, cramers.v = cs$CramersV))
}

# for chi-square
get_chi <- function(x_var, y_var, data, cell) {

  # Convert strings into symbols and save strings
  x_str <- x_var
  y_str <- y_var
  x_var <- sym(x_var)
  y_var <- sym(y_var)
  
  # Chi-square test and Cramer's V
  cs <- chisq.test(pull(data, x_var), pull(data, y_var)) %>%
    tidy() %>%
    mutate(CramersV = lsr::cramersV(
      pull(data, x_var), pull(data, y_var)))
  
  # Return objects
  return(tibble(IV = x_str, DV = y_str, stat = cs$statistic, 
                p = cs$p.value, cramers.v = cs$CramersV,
                method = "chisq", cell = cell))
}

# for fisher's exact
get_fisher <- function(x_var, y_var, data, cell) {

  # Convert strings into symbols and save strings
  x_str <- x_var
  y_str <- y_var
  x_var <- sym(x_var)
  y_var <- sym(y_var)
  
  # Fisher's exact test
  ft <- fisher.test(pull(data, x_var), pull(data, y_var)) %>%
    tidy()
  
  # Return objects
  return(tibble(IV = x_str, DV = y_str, stat = NULL, p = ft$p.value, 
                cramers.v = NULL, method = "fisher", cell = cell))
}

get_stat_wrapper <- function(x_var, y_var, data) {
  num_levels <- data %>%
    select(admin_id, x_var, y_var) %>%
    drop_na() %>%
    pivot_longer(-admin_id) %>%
    group_by(name) %>%
    summarize(n = n_distinct(value)) %>%
    pull(n)
  if(1 %in% num_levels) {
    cell_size_str = NA
    get_fisher(x_var, y_var, data, cell_size_str)
  } else {
    tmp <- chisq.test(pull(data, x_var), pull(data, y_var))
    cell_size_str <- paste0(round(tmp$expected, 2), collapse = " ")
    if(any(tmp$expected < 5)) { # if expected count is less than 5, user fisher's test
      get_fisher(x_var, y_var, data, cell_size_str)
    } else {
      get_chi(x_var, y_var, data, cell_size_str)
    }
  }
}
```

# Attrition Rate and Comparisons

```{r}
# Attrition Rate = (Completed Pre Only) / (Completed Pre) = 26/87 = 29.9%
percent(nrow(df.pre.only) / (nrow(df) + nrow(df.pre.only)), accuracy = 0.1)

# Compare Pre-Test Demographics Between Attritors (pre.a) and Completors (pre.c)
pre.a <- df.pre.only %>% select(starts_with("analysis_pre_"))
pre.c <- df %>% select(starts_with("analysis_pre_"))
map2_dbl(
  names(pre.a), names(pre.c),
  ~ fisher.test(table(pre.a[.x]), table(pre.c[.y]))$p.value
)
```

# Bivariate Tests With Original Data

```{r warning=F, error=F, message=F}
# Select Variables
ivs <- rep(df %>% select(starts_with("analysis_pre_")) %>% names(),
  times = df %>% select(starts_with("analysis_post_")) %>% names() %>% length()
) # number of dvs
dvs <- rep(df %>% select(starts_with("analysis_post_")) %>% names(),
  each = df %>% select(starts_with("analysis_pre_")) %>% names() %>% length()
) # number of ivs

# Fit Models
all_tests <- map2_dfr(ivs, dvs, get_stat_wrapper, data = df)

# Print Results
options(knitr.kable.NA = "")
all_tests %>%
  mutate(p.bonferroni = p.adjust(all_tests$p, method = "bonferroni")) %>%
  arrange(p.bonferroni, p) %>%
  mutate(across(starts_with("p"), ~ ifelse(. < .05, paste0(round(., 3), "*"), round(., 3)))) %>%
  rowid_to_column(var = "#") %>%
  kbl(digits = 3) %>%
  kable_paper()
```

## Significant Result

```{r warning=F, error=F, message=F}
# Check Assumptions for Chi-Square Test
any(chisq.test(df$analysis_pre_01_vol, df$analysis_post_10_effective)$expected < 5)

# Assumption Not Met, Therefore Use Fisher's Test
# See above table for Bonferroni adjusted p-value
fisher.test(df$analysis_pre_01_vol, df$analysis_post_10_effective)

# Cramer's V
lsr::cramersV(df$analysis_pre_01_vol, df$analysis_post_10_effective) # Cramer's V = 0.63

# Plot Significant Result
easy_chi("analysis_pre_01_vol", "analysis_post_10_effective", df)
```

# Bivariate Tests With Imputed Data

```{r warning=F, error=F, message=F}
# Select Variables
ivs <- rep(df %>% select(starts_with("analysis_pre_")) %>% names(), 
           times = df %>% select(starts_with("analysis_post_")) %>% names() %>% length()) # number of dvs
dvs <- rep(df %>% select(starts_with("analysis_post_")) %>% names(),
           each = df %>% select(starts_with("analysis_pre_")) %>% names() %>% length) # number of ivs

# Function to fit chi-square models and pool results
# Code from https://nerler.github.io/EP16_Multiple_Imputation/slide/08_analysis_and_pooling.pdf
get_chisq_pooled <- function(iv, dv) {
  m <- eval(parse(text = glue("with(data = df.imputed, exp = chisq.test(table({iv} , {dv})))")))
  dfs <- m$analyses[[1]]$parameter
  stat <- sapply(m$analyses, "[[", 'statistic')
  m.result <- miceadds::micombine.chisquare(dk = stat, df = dfs, display = F)
  p <- m.result[2]
  return(tibble(IV = iv, DV = dv, p = p))
}

# Fit Models
chisq_ps_imputed <- map2_dfr(ivs, dvs, get_chisq_pooled)

# Print Results
chisq_ps_imputed %>%
  rowid_to_column(var = "#") %>%
  mutate(p.bonferroni = p.adjust(chisq_ps_imputed$p, method = "bonferroni")) %>%
  arrange(p.bonferroni, p) %>%
  mutate(across(starts_with("p"), ~ifelse(. < .05, paste0(round(., 2), "*"), round(., 2)))) %>%
  kbl(digits = 2) %>%
  kable_paper()
```

# Recoded Open-Ended Responses

```{r}
get_openended <- function(var) {
  df %>%
    count({{ var }}) %>%
    mutate(pct = n/sum(n)) %>%
    arrange(desc(pct)) %>%
    mutate(pct_f = percent(pct, accuracy = .1)) %>%
    kbl(digit = 3) %>%
    kable_paper()
}
get_openended(openended_pre_01_concerns)
get_openended(openended_post_01_negatives)
get_openended(openended_post_02_well)
get_openended(openended_post_03_notwell)
```